Based on the quantitative and qualitative data presented in Table \ref{tab:probe_comparison}, the following conclusions can be drawn:

The \textbf{SEGGER J-Link}, while representing the industry standard for professional embedded development, was found to be the least suitable for the course environment due to its prohibitive cost (800 €) and the added complexity of its proprietary software stack. Conversely, the \textbf{pico-debug (Core)} implementation represents the most efficient solution in terms of cost and physical footprint. However, its low maintenance score and the archived status of its RP2040 firmware introduce a significant risk for a curriculum intended to remain operational for several years.

The \textbf{Second Pico} implementation achieved the highest total qualitative score (12.3/15). This methodology offers a "middle ground" that provides the high-level maintenance support of an official Raspberry Pi product while remaining affordable for large-scale classroom deployment. Its ability to provide true dual-core debugging (SMP support) without the maintenance risks associated with virtual probes makes it the most robust choice for a multi-year educational program.

Consequently, while \textit{pico-debug} serves as an excellent zero-cost entry point for individual students, the \textbf{Second Board (Picoprobe)} implementation is recommended as the primary debugging methodology for the Computer Systems course. When integrated with the custom IDE dashboard and 3D-printed casing, it provides the most balanced, reliable, and "snappy" user experience for incoming students.