\section{Virtual Machine VS Container}
To consistently replicate the software environment needed for debugging on Diverse computing infrastructures, two approaches were considered: Virtual Machine and Containers.

To ensure objective evaluations, the underlying operating system was standardized as Debian 12 (Bookworm) across both the containerized and virtualized implementations. The dsitribution was chosen for its lightweight footprint and stability and proven usage in the Computer System course through Xubuntu, a Debian-based operating system[1].

A functional debugging set up requires a software stack on top of the operating system to support it. The software stack used can be divided into four categories:

\begin{itemize}
    \item \textbf{Compiler Toolchain:} The suite of softwares needed to compile C/C++ source code into machine-readable binaries (e.g., UF2, ELF). The compile toolchain's components used for this thesis are \texttt{gcc-arm-none-eabi} cross-compilation infrastructure, which includes \texttt{libnewlib-arm-none-eabi} and \texttt{libnewlib-dev}, used as they provide the standard C library implementation, including memory management, mathematical functions, and system call stubs required for embedded I/O operations. In assistance with the compilation infrastruction, \texttt{CMake} and \texttt{Ninja} serve as the build system and generator, respectively. Together, they provide a versatile and reliable framework for developement of the Computer System course's project and more advanced embedded systems projects.
    
    \item \textbf{Middleware and SDKs:} To unlock all the features supported by the Raspberry Pi Pico platform and complex task management features, the environments also came prepared with Pico and FreeRTOS SDK. These provide the essential hardware abstraction and real-time scheduling capabilities required to support the course's embedded systems projects.
    
    \item \textbf{Development Interface:} To provide a familiar but versatile interface for project development, VS Code was selected for its ecosystem and flexibility, serving as the standardized IDE for the course. Along side with the main IDE, helper extensions are installed to provide effective interaction between the IDE and the development platform:
    
    \begin{itemize}
    \item \textbf{Cortex-Debug:} Facilitates GDB-based debugging, enabling step-through execution and register-level inspection.
    \item \textbf{Raspberry Pi Pico:} Provides essential SDK management and project configuration utilities.
    \item \textbf{Pico-Debug-Dashboard:} A custom-made extension with the intention of unifying how VS Code invokes the compiler toolchain and speaks with the hardware. Supports features such as:
    
    \begin{itemize}
    \item \textbf{Automated Project Bootstrapping:} Instant initialization of standardized project structures, including \texttt{CMake}, \texttt{FreeRTOS}, and debug configurations.
    \item \textbf{Dynamic Build Configuration:} Real-time switching between RP2040 and RP2350 target architectures and debug modes (1-core vs. 2-core SMP) via the status bar.
    \item \textbf{Integrated Build Orchestration:} Single-click access to \texttt{Ninja} build processes, \texttt{GDB} flash deployment, and \texttt{Picotool} operations.
    \item \textbf{Hardware Lifecycle Management:} Centralized controls for hardware resets, BOOTSEL re-entry, and low-level firmware injection (\textit{pico-debug} RAM-loading).
    \item \textbf{System Telemetry:} Real-time visual monitoring of USB bus states and debugger connectivity directly from the IDE status bar.
    \end{itemize}

    \end{itemize}
    As graphical interface are not natively supported by Docker containers, code-server is instead used in place of the original VS Code to provide a browser-based equivalent to VS Code, ensuring a consistent user experience while maintaining the isolation and reproducibility of containerized environments.
    
    \item \textbf{Connectivity and Drivers:} The software components that bridge the host machine with the microcontroller. This includes OpenOCD as the GDB to CMSIS-DAP server, and USB/HID drivers (such as \texttt{libhidapi} and \texttt{libusb}) necessary to maintain communication with the debug probe.
    
    As the virtualization layers lack native access to usb devices, the accessibility must be managed by the virtualization Infrastructure itself. While VM softwares support USB passthrough natively, for Docker containers, a piece of software called \texttt{usbipd} must be installed on the host device. \texttt{usbipd} is a terminal software with the purpose of redirecting the physical USB devices from a Windows host to a guest virtual environment or the Windows Subsystem for Linux (WSL2). It functions by encapsulating USB traffic within TCP/IP packets, thereby overcoming the physical device isolation inherent in hypervisor-based architectures and allowing targeted devices to be seamlessly mounted within the execution environment. To provide easy usage of \texttt{usbipd}, a specialized script file is made for the user to initialize and manage the tool.
\end{itemize}

All software components, including toolchain binaries, SDK releases, and debugging middleware, were configured according to their respective official documentation and established best practices and efforts were made to harmonize the configuration parameters across both the virtual machine and containerized environments. This is to ensure that any observed differences in performance or usability could be attributed strictly to the underlying virtualization architecture.
%[1] https://lovelace.oulu.fi/tietokonej%C3%A4rjestelm%C3%A4t/tietokonej%C3%A4rjestelm%C3%A4t/tkj-virtual-machine-download-and-history/
\section{Metrics description}

Since there are several candidates for the implementation of the probe, we will define a set of metrics to choose the best one. Our set will include both quantitative and qualitative metrics, chosen based on the goal of the project, which is to implement a solution used throughout the educational process. While the values of the quantitative metrics will come directly from the tests and measurements, for qualitative metrics each of the team members will decide on a score from 1 to 5 and the average of those will be used in further considerations.

\subsection{Quantitative metrics}

The first metric considered is the \textbf{price of the hardware}. That is, the price of the plastic or other materials used for the casing, which would fit the TKJHAT and the probe together, as well as the price of the additional microcontrollers, boards, and wires used in the design. Since tens of debugger units are required by the course, the price of each one is a significant detail. 

Next factor we will consider is the \textbf{time to produce} the probe, which includes the printing time of the case for the given probe and the time for soldering and other assembling activities.   

The third quantative metric is the \textbf{physical size} of the final result. That is the size of the casing together with the TKJHAT and the probe. We will also consider the virtual size as a separate metric. It can be refered as \textbf{debug overhead} metric and it estimates the load on the CPU, memory and time consumption added by the debugger. While it might be one the most crucial metrics when working on big and complicated projects, it only takes secondary position in our research, since most of the projects at the entry level course are quite simple and compact.



    Slowdown Factor: Often measured in production debugging, this is the ratio of execution time with the debugger active vs. inactive (e.g., a "10x–100x slowdown" in intensive debugging scenarios).

    Memory Access Speed: The speed at which a debugger reads/writes data to the target system, crucial for embedded processors.

    Breakpoint Latency: Time taken to halt execution when a breakpoint is hit. 

\subsection{Qualitative metrics}

Qualitative metrics play a more important role in our project. Among these is how \textbf{well-maintained} the software to be used is: how often it is updated, if at all; how many people are behind it and how reliable it is in the long run. It is important since the outcome of our work is planned to be in use for a long time, about seven years, and it is likely, that there will be new bugs found and changes made in the upstream software.

One of the most crucial criteria is how \textbf{beginner-friendly} the setup and the usage of the probe is, how likely are the people, working with the embedded systems for the first time, to make a mistake, how easy it is to explain the workflow.

The general \textbf{comfort of use} is another metric we will consider. The hardware setup will affect our ratings as well as the response speed. We consider speed here, in the qualitative category, as exact miliseconds are not so important to us as the general feel and the smootheness of the process.

As a separate and final metric we will consider the \textbf{limitations} of each approach, which might not show in the rest of the metrics.


\section{Probe implementations}
\subsection{Second RP2040/RP2350 implementation}

Following the intruction provided by Raspberry Pi \cite{pico_getting_started}, the second RP2040/RP2350 device, now referred to as the probe, is flashed with the \texttt{debugprobe.uf2}. Physical connections between the probe and the targeted device are made according to the guide (see Page \pageref{box:picoprobe-wiring}). Once the probe is connected to the host machine via USB, the user can start the debugging session through VS Code \cite{pico_getting_started} or initializing the installed GDB and OpenOCD through command-line with commands:

\$ sudo openocd -f interface/cmsis-dap.cfg -f target/rp2040.cfg -c "adapter speed 5000"

For a RP2350-based device use -f target/rp2350.cfg instead.

Then a specialized version of GDB is started with 

\$ gdb-multiarch

Finally, the GDB client connects to the \texttt{OpenOCD} server via \texttt{target remote localhost:3333}, establishing the data link required to execute, inspect, and monitor the target device's firmware in real-time.

\subsection{Second core implementation - pico-debug}

The approach where one core is used to debug another one was implemented and shared by majbthrd on Github. He developed two projects: "pico-debug" for RP2040 \cite{pico_debug}, and "pico2-debug" for RP2350 boards \cite{pico2_debug}. To setup the debugging session there are ready .uf2 files provided in the repositories. Pico-debug can be placed in RAM or cache memory of the board, and each option has its own .uf2 file. We are going with a more common and easier way to use it, placing it in the RAM, so the whole cahce is availiable for the code of the program that is being debugged. To do so, we take the pico-debug-gimmecache.uf2 file from the releases page of the repository and flush it onto the board. 

Now the board shows as a CMSIS-DAP probe in the list of devices, and we can proceed to connect to it through the GDB and OpenOCD. The OpenOCD, however, has to be built by the user, since the newer versions of its Raspberry Pi's fork are not compatible with pico-debug. The detailed instructions are provided in the repository's "howto" section \cite{pico_debug}. 

\subsection{Jlink}
